---
title: "Question2"
author: "Sheetal"
date: "7/6/2018"
output: html_document
---

```{r setup, include=FALSE}
library("xlsx")
setwd("/Users/sheetalgolecha/Documents/DMG1/individual assignment")
getwd()
```

## R Markdown

Reading the file and creating the PCA
```{r cars}
input<-read.csv("winedata.csv", header=TRUE)
dim(input)

mydata <- input[1:178,2:14]
head(mydata)

pcaObj <- princomp(mydata, cor = TRUE, scores = TRUE, covmat = NULL)   ### scored= true mean new data set will be created covmat= covariance matrix

summary(pcaObj)
plot(pcaObj)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
loadings(pcaObj)   ##
pcaObj$loadings
```

Constructing the biplot to look at the polarity of columns and check pc scores



```{r}
biplot(pcaObj)
biplot(pcaObj$scores[,1:2],pcaObj$loading[,1:2]) ## if you didn't want rescaling of the axis graph between comp1 and comp2

pc_scores <- pcaObj$scores   # on the object all the scores
pc_scores

scale(mydata)    ### take the standardize matrix and matrix multiple with locading to get new data.
scale(mydata) %*% pcaObj$loadings ## this is same as pcaObj$scores

```

Scaling the data using two componenets of PCA and calculating the rmse of the data


```{r}
scaled_data <- scale(mydata)
rmse <- matrix(nrow=13,ncol=1)
for (i in 1:13) {
  scaled_data_recovered <- pc_scores[,1:i] %*% t(pcaObj$loadings[,1:i]) ## if compressed using two components
  scaled_data_recovered
  rmse[i] <- sum((scaled_data_recovered - scaled_data)^2) ## errors
}
rmse
```

Visualizing how the data is spread among all the clusters.

```{r}
pcaObj$scores

plot(pcaObj$scores[,1], pcaObj$scores[,2], col="blue", xlab="Component 1", ylab="Component 2")


```

Looking at the screeplot to define the cluster type and applying K mean algorithm to apply it on the cluster.

```{r}
normalized_data=scale(scaled_data_recovered)
Cluster_Variability <- matrix(nrow=5, ncol=1)
for (i in 1:5) Cluster_Variability[i] <- kmeans(normalized_data,centers=i, nstart=4)$tot.withinss
plot(1:5, Cluster_Variability, type="b", xlab="Number of clusters", ylab="Within groups sum of squares") ## Elbow curve or Scree plot

fit <- kmeans(normalized_data, centers=2, iter.max=10, nstart=4)
NormDatasWithClusterMembership <- cbind(input, fit$cluster)
write.xlsx(NormDatasWithClusterMembership,file="winedata_cluster.xlsx")
```


Applying the Kmean on the complete set.


```{r}
normalized_data_1=scale(mydata)
Cluster_Variability <- matrix(nrow=5, ncol=1)
for (i in 1:5) Cluster_Variability[i] <- kmeans(normalized_data_1,centers=i, nstart=4)$tot.withinss
plot(1:5, Cluster_Variability, type="b", xlab="Number of clusters", ylab="Within groups sum of squares") ## Elbow curve or Scree plot

fit <- kmeans(normalized_data_1, centers=2, iter.max=10, nstart=4)
NormDatasWithClusterMembership_1 <- cbind(input, fit$cluster)
write.xlsx(NormDatasWithClusterMembership_1,file="winedata_cluster_real_ds.xlsx")
```

